---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ovbsa

<!-- badges: start -->
<!-- badges: end -->

The goal of `ovbsa` is to conduct sensitivity analysis of omitted variable bias in linear econometrics models.

## Installation

You can install the development version of ovbsa from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("dbasu-umass/ovbsa")
```

## Example 1

This is a basic example which shows you how to find the bias-adjusted estimate, the bias-adjusted standard error and the bias-adjusted confidence interval in a linear regression model. The user needs to choose a benchmark covariate, values of the sensitivity parameters `kD` and `kY` and the significance level `alpha` for testing the null hypothesis that the treatment effect is zero. 

Let us first load the relevant libraries.

```{r}
library(ovbsa)
library(sensemakr)

```

In this example we will use use the data set `darfur` from the package `sensemakr`, which studies the effect of exposure to violence on attitudes towards peace. Here we choose `kD=3`, `kY=3` and `alpha=0.05`

```{r example1}
## basic example code
analysis1 <- ovbsa::bsal(
  kd=3,ky=3,alpha=0.05,data=darfur,
  outcome = "peacefactor",
  treatment = "directlyharmed",
  bnch_reg = "female",
  other_reg = c("village","age","farmer_dar","herder_dar",
                "pastvoted","hhsize_darfur")
)
```

Now, let us see the results.

```{r}
(analysis1)
```

## Example 2

Continuing with the previous example, we will now compute the probability that taking account of omitted variable bias will overturn the conclusion of the study. Here we choose `alpha=0.05`.

First, total R-squared based analysis:
```{r}
# total r2-based analysis
res1 <- ovbsa::saltr2(
  alpha=0.05,data = darfur, outcome = "peacefactor",
  treatment = "directlyharmed", bnch_reg = "female",
  other_reg = c("village","age","farmer_dar","herder_dar",
                "pastvoted","hhsize_darfur"),
  N = 500, k_kd=1, k_ky=1
)
# probability
(res1$frac_prob_wt)
```

Second: partial R-squared based analysis without conditioning on the treatment variable:
```{r}
res2 <- ovbsa::salpr2ncd(
  alpha=0.05,data = darfur, outcome = "peacefactor",
  treatment = "directlyharmed", bnch_reg = "female",
  other_reg = c("village","age","farmer_dar","herder_dar",
                "pastvoted","hhsize_darfur"),
  N = 500, k_kd=1, k_ky=1
)
# probability
(res2$frac_prob_wt)

```

Finally, partial R-squared based analysis with conditioning on the treatment variable:
```{r}
res3 <- ovbsa::salpr2cd(
  alpha=0.05,data = darfur, outcome = "peacefactor",
  treatment = "directlyharmed", bnch_reg = "female",
  other_reg = c("village","age","farmer_dar","herder_dar",
                "pastvoted","hhsize_darfur"),
  N = 500, k_kd=1, k_ky=1
)
# probability
(res3$frac_prob_wt)

```



